### 2024-01-01 chapters/background.tex
Polish background

Instead, \textcite{jaynes_information_1957,jaynes_prior_1968} argued that the information entropy of a continuous distribution should be defined as the limiting density of increasingly dense discrete distributions, $H_{N\rightarrow\infty}(X)$. This argument leads to,
%
\begin{equation} \label{eq:corrected-differential-entropy-jaynes}
    H_{N\rightarrow\infty}(X) \equiv \lim_{N\rightarrow\infty} [H_{N}(X) - \log N] = - \int_{\mathcal{X}} p_X(x) \log \left(\frac{p_X(x)}{q_X(x)}\right) \, dx \enspace ,
\end{equation}
%
where $H_N(X)$ is the limiting density for discrete points, $q_X(x)$ is a uniform density over the quantization of the continuous space, and we have subtracted a $\log N$ term that would go to infinity in the limit of infinite points.
By doing this, the information entropy becomes positive, dimensionally correct, and invariant under a change of variables.%
\footnote{\label{fn:change-of-variables-invariance-of-corrected-differential-entropy}
    Under a change of independent variable from $x$ to $y(x)$, we have that
    \begin{align*}
        \widetilde{w}(y)\,dy = w(x)\,dx \enspace , \quad
        \widetilde{q}(y)\,dy = q_X(x)\,dx \enspace .
    \end{align*}
    Plugging this into the differential entropy \cref{eq:corrected-differential-entropy-jaynes} we arrive at the original expression but now with $y$ as the independent variable,
    \begin{equation*}
        H_{N\rightarrow\infty}(y(x)) =  = - \int_{\mathcal{Y}} \widetilde{w}(y)\,\frac{dy}{dx} \log \left( \frac{\widetilde{w}(y)\frac{dy}{dx}}{\widetilde{q}(y)\frac{dy}{dx}} \right) \, dx = - \int_{\mathcal{Y}} w(y) \log \left( \frac{w(y)}{q(y)} \right) \, dy \enspace .
    \end{equation*}
}

@@

Returning to Jaynes' argument and \cref{eq:corrected-differential-entropy-jaynes}, we can interpret this to say that the information entropy of a continuous random variable $X$ should be defined as the expected difference in information entropy between its density and a uniform density. 
Identifying the Kullback-Leibler divergence as a measure of this relative information entropy provides a natural point of convergence for this section since it is equivalently defined for both discrete and continuous random variables, contrary to the differential entropy. 
In this view, information entropy is a quantity that distills the distribution of a random variable into a single number that describes the diversity of the potential outcomes of the random variable.

@@

\section{Out-of-distribution detection} \label{sec:out-of-distribution-detection}
% \lesstodo[inline]{Maybe this section should come after the VAE background?}
% 
% OUTLINE: C.f. https://openaccess.thecvf.com/content/CVPR2023W/VAND/papers/Graham_Denoising_Diffusion_Models_for_Out-of-Distribution_Detection_CVPRW_2023_paper.pdf
% - Formalize the problem. https://arxiv.org/pdf/2302.10326.pdf
%   - Assume we have a dataset $\mathcal{D}_{\text{train}}$ of in-distribution data and we would like to build a model that can be used to assess whether a certain data point $\xb$ is sampled from $\mathcal{D}_{\text{train}}$ or not. Given a test point $\xb$, the model outputs a score $s(\xb)$ that indicates whether $\xb$ is in-distribution or out-of-distribution. 
%   - In the context of neural networks, this problem dates back to at least 1994 \parencite{bishop_novelty_1994}. 
%
% - Categorisation into supervised and unsupervised. 
%   - Supervised: Whether p(y|x) should be trusted for a given x. 
%   - Unsupervised: Whether x should be trusted and used at all. 
%
In \cref{subsec: overview-of-out-of-distribution-detection} we introduce out-of-distribution (OOD) detection and connect it to the concept of uncertainty. 
Since the related work sections of the papers in \cref{chp:paper-hierarchical,chp:paper-modelagnostic} are relatively short, we also provide a concise review of the literature on out-of-distribution detection in \cref{subsec: supervised-out-of-distribution-detection, subsec: unsupervised-out-of-distribution-detection}.

>>>>

Instead, \textcite{jaynes_information_1957,jaynes_prior_1968} argued that the information entropy of a continuous distribution should be defined as the limiting density of increasingly dense discrete distributions, $H_{N\rightarrow\infty}(X)$. This argument leads to,
%
\begin{equation} \label{eq:corrected-differential-entropy-jaynes}
    H_{N\rightarrow\infty}(X) \equiv \lim_{N\rightarrow\infty} [H_{N}(X) - \log N] = - \int_{\mathcal{X}} p_X(x) \log \left(\frac{p_X(x)}{q_X(x)}\right) \, dx \enspace ,
\end{equation}
%
where $H_N(X)$ is the limiting density for discrete points, $q_X(x)$ is a uniform density over the quantization of the continuous space, and we have subtracted a $\log N$ term that would go to infinity in the limit of infinite points.
By doing this, the information entropy becomes positive, dimensionally correct, and invariant under a change of variables.%
\footnote{\label{fn:change-of-variables-invariance-of-corrected-differential-entropy}
    Under a change of independent variable from $x$ to $y(x)$, we have that
    \begin{align*}
        \widetilde{w}(y)\,dy = w(x)\,dx \enspace , \quad
        \widetilde{q}(y)\,dy = q_X(x)\,dx \enspace .
    \end{align*}
    Plugging this into \cref{eq:corrected-differential-entropy-jaynes} we arrive at the original expression but now with $y$ as the independent variable,
    \begin{equation*}
        H_{N\rightarrow\infty}(y(x)) =  = - \int_{\mathcal{Y}} \widetilde{w}(y)\,\frac{dy}{dx} \log \left( \frac{\widetilde{w}(y)\frac{dy}{dx}}{\widetilde{q}(y)\frac{dy}{dx}} \right) \, dx = - \int_{\mathcal{Y}} w(y) \log \left( \frac{w(y)}{q(y)} \right) \, dy \enspace .
    \end{equation*}
}

@@

Returning to Jaynes' argument and \cref{eq:corrected-differential-entropy-jaynes}, we can interpret this to say that the information entropy of a continuous random variable $X$ should be defined as the expected difference in information entropy between its density and a uniform density. 
Identifying the Kullback-Leibler divergence as a measure of this relative information entropy provides a natural point of convergence for this section since it is equivalently defined for both discrete and continuous random variables, contrary to the differential entropy. 
In this view, information entropy is a quantity that distills the distribution of a random variable into a single number describing the diversity of the potential outcomes of the random variable.

@@

\section{Out-of-distribution detection} \label{sec:out-of-distribution-detection}
% \lesstodo[inline]{Maybe this section should come after the VAE background?}
% 
% OUTLINE: C.f. https://openaccess.thecvf.com/content/CVPR2023W/VAND/papers/Graham_Denoising_Diffusion_Models_for_Out-of-Distribution_Detection_CVPRW_2023_paper.pdf
% - Formalize the problem. https://arxiv.org/pdf/2302.10326.pdf
%   - Assume we have a dataset $\mathcal{D}_{\text{train}}$ of in-distribution data and we would like to build a model that can be used to assess whether a certain data point $\xb$ is sampled from $\mathcal{D}_{\text{train}}$ or not. Given a test point $\xb$, the model outputs a score $s(\xb)$ that indicates whether $\xb$ is in-distribution or out-of-distribution. 
%   - In the context of neural networks, this problem dates back to at least 1994 \parencite{bishop_novelty_1994}. 
%
% - Categorisation into supervised and unsupervised. 
%   - Supervised: Whether p(y|x) should be trusted for a given x. 
%   - Unsupervised: Whether x should be trusted and used at all. 
%
In \cref{subsec: overview-of-out-of-distribution-detection} we introduce out-of-distribution (OOD) detection and connect it to the concept of uncertainty. 
Since the related work sections of the papers in \cref{chp:paper-hierarchical,chp:paper-modelagnostic} are relatively short, we also provide a concise review of the literature on out-of-distribution detection in \cref{subsec: supervised-out-of-distribution-detection} and \ref{subsec: unsupervised-out-of-distribution-detection}.
